---
title: "Section 5"
format: revealjs
editor: visual
---

## Section in 30ish Second

-   Cover some common mistakes in PS1 and PS2

-   Questions

-   Regression Practice

    -   Available on bCourses as "Section 5 Activity"

## Common Mistakes

I only grade one submission for fairness reasons. Here are the primary reasons students have lost points so far:

-   Forbidden Packages

-   Not answering all parts of a question

-   Incorrectly updating rows in a data frame

-   Incorrect loops

## Forbidden Packages

-   Points have to be taken off for using functions from packages not allowed on the assignment. **Only use packages explicitly allowed on the assignment.**

    -   Functions from the following packages are always allowed `stats, graphics, grDevices, utils, datasets, methods, base.`

    -   Functions from `ggplot2` are always allowed unless noted otherwise.

-   If you're not sure if you can use a function or a package, please ask me or post to Ed Discussion.

## All Parts of the Question Need to be Answered

-   I cannot give you points if there is no answer to a part of a question.

-   Generally speaking, if a question says "Comment" or an a synonym of comment somewhere you need to write some English in addition to the math.

-   If a question says "Report" you need to output an answer.

## Updating Rows in a Data Frame

-   The usual method is `data$col[i] = update`.

    -   `data[["col"]][i] = update` also works and is more general

-   If you update a vector outside of a data frame, it will not update the data frame column.

-   Generally speaking, we should always update the column the question asks us to update.

## Updating Rows in a Data Frame

```{r, echo = T}
u = 1:6
v = 1:6
dat = data.frame(u,v)
head(dat)
```

```{r, echo = T}
v[1] = 10 
print(v)
head(dat)
```

Why does R not update the data frame column?

## Incorrect Loops (Part 1)

-   Not actually looping through data

```{r, echo = T}
t = c(1,2,3)
# length(t) == 3 
for(i in length(t)){
  print(t)
}
```

-   Trying to pass more than one value to a single element of a vector

```{r, echo = T, warning = T}
x = vector(mode = "logical", length = 1)
x[1] = c(1,2,3,4,5)
print(x)
```

## Incorrect Loops (Part 2)

-   Not sampling the entire vector with replacement

```{r, echo = T}
y = vector(mode = "logical", length = 1)
y = mean(sample(1:100, 5, replace = T))
print(y)

```

-   Hardcoding a variable in the function

```{r, echo=T}

same_mean_3_times = function(z){
  w = vector(mode = "logical", length = 3)
  for(i in 1:3){
    w[i] = mean(sample(1:100, replace = F))
  }
  return(w)
}
same_mean_3_times(z = 1:5) ## want to get the output c(3,3,3)
```

## Incorrect Loops (Part 3)

The following do not lose points but are strongly discouraged. See the Loop Speed handout on bCourses on Ed Discussion for why.

### c() 

```{r, echo = T, eval = F}
v = c()
x = 1:100
for(i in 1:N){
  v = c(v, mean(x))
}
```

### append()

```{r, echo = T, eval = F}
v = c()
for(i in 1:100000){
  v = append(v, mean(sample(x))
}
```

## Things that will lose points on future assignments

-   Printing the entirety of a data frame or vector that is large because it makes it hard for me to find your answers

    -   use `head(dat)` or `vec[1:6]`

-   Leaving in the extraneous template code and words

    -   e.g. *"Add text/explanation..."*
    -   Make sure to delete this before submission

## Questions from Lecture

-   Note we've done multiplication on vectors in R like this

```{r, echo = T}
a = c(1,2,3)
b = c(2,3,4)
a*b
```

In Lecture it was mentioned that the inner product is

```{r, echo =T}
a%*%b
```

The difference between these is that the first computes "element-wise" multiplication. The second in the inner product.

-   Other questions about the course material?

## Section 5 Activity

-   The activity for today is on bCourses.

    -   It is drawn from your textbook with some minor modifications.

-   If/When you finish the "Main" section, throw up your hand(s).

-   The "Bonus" section provides some additional practice if you finish early.
