---
title: "Section 2 R Activity"
format: 
  html:
    toc: true
    toc-title: Outline
    toc-location: left
editor: visual
---

## How should we think about randomness?

```{r}

set.seed(123)
num_vec = 1:5
num_vec[c(5,3,1,2,4)]
num_vec[c(1,3,4,2,5)]
```

On your own computer, try to make a different order than the ones on the screen.

### How many possible ordering are there are 5 numbers?

```{r}
factorial(5)
```

### What are we doing here?

SAMPLING!

```{r}
sample(1:5, 5)
```

## Section 101

```{r, echo=FALSE, message =FALSE}
sections = read.csv("https://tinyurl.com/4uarvs4c")
section101 = subset(sections, section == 101)
section102 = subset(sections, section == 102)
```

```{r}
head(section101)
```

### Let's make some random groups

How many students do we have

```{r}
nrow(section101)
```

### Let's make 9 groups

```{r}
group = 1:9
section101$group = sample(group, nrow(section101), replace = TRUE)
section101[order(section101$group),]
```

### We have a problem when we do this. Why?

### An alternative

```{r}
group2 = c(rep(1:9, each = 3))
section101$group = sample(group2, 
                          nrow(section101), 
                          replace = FALSE)
section101[order(section101$group),]
```

## Section 102

```{r}
head(section102)
```

### Let's make some random groups

How many students do we have?

```{r}
nrow(section102)
```

### Let's make 9 groups

```{r}
group = 1:9
section102$group = sample(group, nrow(section101), 
                          replace = TRUE)
section102[order(section102$group),]
```

### We have a problem when we do this. Why?

### An alternative

```{r}
group2 = c(rep(1:9, each = 3))
section102$group = sample(group2, nrow(section101), replace = FALSE)
section102[order(section102$group),]
```
