---
title: "PS132B Section 1"
format: beamer
editor: visual
---

## Section in 30ish seconds

-   Fill out the Office Hour form on Ed Discussion

-   Participation Activity

    -   Reply to Ed Discussion there is a thread called "Section 1 Activity" with the name of a song that you think everyone should listen to. Include the artist name as well

-   Learn some other peoples' names

-   Go over the survey results

    -   If you put "None at all" or "A little" to coding experience you're the modal respondent

## Office Hours

I am deciding when to hold my office hours between the following day/times:

-   M: 10-12

-   T: 10-12

-   W: 2-4pm

-   H: 10-12

-   F: 2-4

Please fill out the form linked on Ed Discussion

## R Installation

The R and RStudio installation instructions are on Ed Discussion

Note: You might need to install tinytex. The instructions to do that are in a thread on Ed Discussion

## Make a Friend

```{r, echo = T, eval = F}
makeFriendFunction = function(){
  ## 1. Survey the classroom 
  ## 2. Find at least one person you have never met 
  ## 3. Introduce yourself to that person 
  ## 4a. Find some common shared aspect of the 
  ## human experience
  ## 5. If appropriate, exchange contact information
}

## Loop the function to make three friends 
for(i in 1:3){
  makeFriendFunction()
}
```

## What are the answers for the survey questions?

```{r, echo = T}
## Makes a vector with the c() function and 
## assigns value to x
x = c(1,2,3,4,5,6,7,8)

## prints to the console the "value" of the object x
print(x)

## a way to get only the even values 
## from the vector x
y = x[seq(2,8,2)] 

## prints the value of the vector y 
print(y)
```

## What are the answers for the survey questions?

```{r, echo = T}
set.seed(123)
## Make a vector of the values 1,3,5,7
x = c(1,3,5,7) 

## Make another vector that takes the values of x 
## and adds noise from a standard normal distribution
y = x + rnorm(4) 

## runs a linear regression of y on x
lm(y ~ x)

```

## What are the answers for the survey questions?

```{r, echo = T}
## This function removes the last value of a vector 
## we might call it dropLast instead of f2
f2 = function(x){ 
  if(length(x)== 1){
    return(NULL) 
  }
  x[-length(x)] 
}

x = c(1,2,3,4,5)
print(x)
f2(x)
```

## What should I take away from today?

-   This class succeeds because of you. Everyone in class can learn the material.

-   We want to help you learn the material.

-   If you do not understand R code, you are not alone in class

-   Type examples out. Avoid copy/pastes

## Some Concrete Suggestions to see what happens (Problem 1)

The following are all suggestions to see *what* happens without worrying necessarily about *why* it happens.

Learning a language is as much about breaking things/trying something and seeing what happens as anything else early on

-   Try changing the number of elements in the vector.
    -   What happens if you make x the first 10 numbers?
    -   What happens if you make `x = c("A", "B", "C", "D", "E")`
    -   What happens if you change the last argument in `seq()` to 4
-   Type ?seq into your R console and press enter. What happens?

## Some Concrete Suggestions to see what happens (Problem 2)

-   Swap the order of y and x. Do you get the same answer? Can you think of why you get the result? Write down some guesses.

-   Add `y = c(y, 2,3)` after the the line defining y initially. Run the same `lm()` function. What happens? Can you think of why you get the result? Write down some guesses.

-   Does it matter that x is composed of just whole numbers? Experiment by changing the values to fractions.

## Some Concrete Suggestions to see what happens (Problem 3)

-   Try passing a vector of length 1 to the function. What happens?
    -   e.g. `x = c(5)`
-   Try passing the function a vector of names. What happens? Remember that we need to put "" around them like this
    -   `x = c("Bansak", "Wurster")`
-   Amend line 7 to be `x[-(length(x)-3)]`. What happens if you run the same example as in class? What happens if you make `x=c(1,2)`? Do you have any guesses as to why this happens?
