---
title: "Minimal Poker Pair Classifier"
author: "Alex Stephenson"
date: "2023-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Is it a pair? 

The following is a program that will classify whether a 5 card hand in poker is a pair or not. For this program, we ignore anything about suits.


## Make a deck 

```{r}
deck = rep(1:13, each = 4)
deck
```


## Pair Classifier 
We can combine table with which() and length() to determine if we have just one pair. 
This could be one line in our function, but I'll break it up for teaching clarity

```{r}
isPair = function(hand){
  values = table(hand)
  pairs = which(values == 2)
  numpairs = length(pairs)
  
  if(numpairs == 1){
    return("Pair")
  } else{
    return("Not Pair")
  }
}
```

## Let's apply our test cases 

```{r}
pair = c(1,1,2,3,4)
twopair = c(1,1,2,2,3)
nopair = c(1,2,3,4,5)
nopair2 = c(1,2,11,11,11)
testcases = list(pair, twopair, nopair, nopair2)
for(i in 1:length(testcases)){
  print(isPair(testcases[[i]]))
}
```

Of course we want to do this repeatedly so a loop seems reasonable.

```{r}
## Set the number of runs
set.seed(1234)
N = 1000

## preallocate a vector
pairs = vector(mode = "logical", length = N)
for(i in 1:N){
  pairs[i] = isPair(sample(deck, 5, replace = F))
}
prop.table(table(pairs))
```

It turns out that our estimate after 1000 runs is pretty close to the theoretical result of 42.26%. 