---
title: "Example of Presenting Results"
author: "Kirk Bansak"
date: "January 31, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Summary

Below are some simple results on asylum applications in the European Union.


## The Data

```{r}
dat <- read.csv("europe_asylum_2021.csv")
head(dat)
```

The countries included in the data set include the following (country codes):

```{r}
unique(dat$geo)
```


## Results

In the European Union in 2021, the mean number of asylum applications per country was the following:

```{r}
mean(dat$napps)
```

In the European Union in 2021, the total number of asylum applications per country was the following:

```{r}
sum(dat$napps)
```

In sum, a total of `r sum(dat$napps)` asylum applications were submitted in Europe in 2021.

Here is a histogram of the distribution of applications across the countries:

```{r}
hist(dat$napps, breaks = 15)
```

The end!
